Meta-Analysis of the `Ironic' Effects of Intergroup Contact\thanks{}

### 2021-12-13 manuscript/manuscript.tex
start revision

\documentclass[12pt, letterpaper]{article}

% Font
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{microtype}

% Format
\usepackage[letterpaper, margin = 1in]{geometry}
\setcounter{secnumdepth}{0}
\setlength{\parindent}{0.5in}
\usepackage[compact]{titlesec}
\titleformat*{\section}{\centering\normalfont\normalsize\bfseries}
\titleformat*{\subsection}{\raggedright\normalfont\normalsize\bfseries}
\titleformat*{\subsubsection}{\raggedright\normalfont\normalsize\bfseries\itshape}
\newcommand{\refsection}{\newpage \section{References}}
\usepackage{indentfirst}
\usepackage[hang,flushmargin]{footmisc}
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Header
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{14.5pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{`IRONIC' EFFECTS META-ANALYSIS}
\fancyhead[R]{\thepage}

@@

% Line spacing
\usepackage[doublespacing]{setspace}

% Frontmatter
\title{ Meta-Analysis of the `Ironic' Effects of Intergroup Contact }

@@

\section{Abstract}

\noindent Growing evidence suggests that intergroup contact, psychology's
most-researched paradigm for reducing prejudice, has the `ironic' effect
of reducing support for social change in disadvantaged groups. We
conducted a preregistered meta-analytic test of this effect across 98
studies with 140 samples of 213,085 disadvantaged-group members. As
predicted, intergroup contact was, on average, associated with less
perceived injustice (\(r = -.07\)), collective action (\(r = -.06\)),
and support for reparative policies (\(r = -.07\)). However, these
associations were small, variable, and consistent with alternative
explanations. Across outcomes, 25--36\% of studies found \emph{positive}
associations with intergroup contact. Moderator analyses explained about
a third of the between-sample variance, showing that, on average,
associations were negative only in studies of adults that measured
intergroup contact directly and were strongest in studies that examined
short-term migration or (post-)colonial intergroup relations. We also
found evidence for an alternative explanation for the apparent `ironic'
effects of intergroup contact as, after controlling for the positive
association of negative contact with support for social change, positive
contact was no longer associated with any of the outcomes. We close by
discussing strengths and limitations of the available evidence and by
highlighting open questions about the relationship between intergroup
contact and support for social change in disadvantaged groups.

\textit{Keywords:} intergroup contact, meta-analysis, perceived discrimination, collective
action, policy support

\newpage

\section{Meta-Analysis of the `Ironic' Effects of Intergroup Contact}

It is ironic that psychology's most-researched paradigm for reducing
prejudice---intergroup contact---should hinder, rather than help, social
change. Hundreds of studies have confirmed that intergroup contact is
associated with less negative feelings toward outgroup members (for a
meta-analysis, see Pettigrew \& Tropp, 2006). Most of these studies have
focused on advantaged-group members' contact experiences with
disadvantaged-group members (Tropp \& Pettigrew, 2005), guided by the
idea that reducing advantaged-group members' prejudice will prevent
interpersonal discrimination. Reducing prejudice, however, is rarely
enough to overcome institutional and structural discrimination. Instead,
as the 2020 Black Lives Matter protests have shown, social change often
requires political mobilization by the disadvantaged. Critics have
argued that, by fostering harmony and reducing conflict, intergroup
contact might distract disadvantaged-group members from the injustice
they face and, thereby, undermine their motivation for social change
(Dixon et al., 2005; Reicher, 2007; Wright, 2003). By reducing
disadvantaged-group members' opposition to injustice, intergroup contact
might thus have the `ironic' (Saguy et al., 2009), `sedative' (Çakal et
al., 2011), or `paradoxical' (Dixon et al., 2010) effect of hindering
social change.

@@

Collective action is any action directed at improving the conditions of
the disadvantaged ingroup (Wright et al., 1990). Collective action is
crucial to achieving social change because advantaged-group members
rarely give up their advantages without sustained pressure from
disadvantaged-group members (Blumer, 1958; for a review, see Dixon et
al., 2012). Researchers have argued that intergroup contact could
discourage disadvantaged-group members from engaging in collective
action by diminishing their perceived discrimination (Tropp et al.,
2012), by blurring boundaries between the disadvantaged ingroup and the
advantaged outgroup (Saguy et al., 2009), by reducing negative attitudes
toward the advantaged outgroup (Wright \& Lubensky, 2009), and by
quelling their anger (Hayward et al., 2018; Tausch et al., 2015). To the
extent that it discourages collective action, intergroup contact reduces
support for social change.

@@

As preregistered, we selected studies in three stages. First, we
screened records based on their title, abstract, and keywords. We
refined our coding strategy over three random samples of 100 records
until we achieved acceptable inter-rater agreement (\(\kappa_1 = .56\),
\(\kappa_2 = .60\), \(\kappa_3 = .79\)). We then divided the remaining
records between the two authors. For each record, one of the authors
coded whether the record met the eligibility criteria (yes, maybe, no),
or whether it was a relevant review article. We kept all records coded
as ``yes'' or ``maybe''. Second, both authors reviewed each of the
full-text manuscripts from the previous stage and coded whether any
study or sample in the manuscript fulfilled the preregistered
eligibility criteria (\(\kappa = .75\)). Third, we resolved any
disagreements (by consensus) and excluded ineligible records.

@@

Developing methods to detect and correct for publication bias and other
meta-biases is an active area of research, with no single method
outperforming all others (Carter et al., 2019). Following Carter et
al.'s recommendations, we compared results of several methods to adjust
meta-analytic estimates for publication bias: the PET-PEESE estimator
(Stanley \& Doucouliagos, 2014); the three-parameter selection model
(3PSM; Vevea \& Hedges, 1995); and the \emph{p}-uniform* estimator (van
Aert \& Assen, 2018). In addition, we ran a subgroup analysis that
included only studies that were not published or not conducted with the
intention to examine the effects of intergroup contact on any of the
outcomes---and that, therefore, can be assumed to have not been affected
by publication bias. As these methods do not use Bayesian statistics, we
compared adjusted estimates to unadjusted estimates from a
random-effects meta-analysis estimated with restricted maximum
likelihood estimation using the \emph{metafor} package (Viechtbauer,
2010).

@@

First, the estimated effect sizes for the average associations of
intergroup contact with perceived injustice (\(r = -.07\)), collective
action (\(r = -.06\)), and policy support (\(r = -.07\)) were small.
Across the three outcomes, the 95\% most plausible estimates included
effect sizes between \(r = -.14\) and \(r = .02\). Effect sizes were
thus much smaller than for the association between contact and prejudice
in minority (\(r = -.18\)) and majority (\(r = -.23\)) groups (Tropp \&
Pettigrew, 2005).\footnote{Comparing standardized effect sizes assumes
  that the two outcomes are of equal importance. That is not always the
  case: for example, a drug that reduces mild symptoms by 0.20 standard
  deviations is not better than a drug that reduces deaths by 0.10
  standard deviations.} Small effects can still be important if they
accumulate over time or across people. For example, even a small change
in policy attitudes could, if it affects enough people, sway a tight
election. Cumulative effects, however, should not be assumed without an
empirical or theoretical rationale (Funder \& Ozer, 2019).

@@

Third, the available evidence is almost entirely from cross-sectional,
observational studies and thus consistent with alternative explanations
for the observed associations. Support for social change could reduce
intergroup contact rather than the other way around. For example,
disadvantaged-group members involved in collective action might avoid
forming friendships with advantaged-group members. Alternatively, the
observed associations could be spurious with both being caused by an
unobserved confounder. For example, a disadvantaged-group member's
socioeconomic status might both expose them to more advantaged-group
members and reduce their perception of injustice. Future research should
prioritize longitudinal studies to confirm the direction of the observed
associations and (field) experiments to rule out confounding and other
alternative explanations.\footnote{Published research includes only
  three longitudinal studies (Koschate et al., 2012; Reimer et al.,
  2017; Tropp et al., 2012), all of which conflated within-person change
  and between-person stability (Hamaker et al., 2015), two experimental
  studies with a no-contact control condition (Becker et al., 2013;
  Droogendyk et al., 2016), and two intervention studies (Reimer et al.,
  2021; Shani \& Boehnke, 2017).}

@@

Our research was motivated by the broader question whether intergroup
contact helps or hinders social change. By examining the narrower
question whether intergroup contact diminishes support for social change
in disadvantaged groups, we tested the most prominent hypotheses
explaining how intergroup contact might \emph{hinder} social change. We
did not consider whether intergroup contact might also \emph{help}
social change by reducing prejudice (for critical perspectives, see
Dixon et al., 2012; Jackman, 1994) or by motivating the advantaged to
acknowledge and challenge social inequality (Tropp \& Barlow, 2018).
Further, we did not consider contact between different disadvantaged
groups, which has been found to increase support for social change among
the disadvantaged (Dixon et al., 2015, 2017). Therefore, this
meta-analysis could not provide a complete answer to the question
whether, on balance, intergroup contact helps or hinders social change.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-3201}{}%
*Albzour, M., Penic, S., Nasser, R., \& Green, E. G. T. (2019). Support
for ``normalization'' of relations between Palestinians and Israelis,
and how it relates to contact and resistance in the West Bank.
\emph{Journal of Social and Political Psychology}, \emph{7}(2),
978--996. \url{https://doi.org/10.5964/jspp.v7i2.877}

\leavevmode\hypertarget{ref-2023}{}%
*Bagci, S. C. (2014). \emph{Cross-ethnic friendships in multiethnic
educational settings: Consequences for psychological, academic and
motivational outcomes among young adolescents} {[}Doctoral dissertation,
Goldsmiths, University of London{]}.
\url{https://doi.org/10.25602/GOLD.00011030}

\leavevmode\hypertarget{ref-2395}{}%
*Bagci, S. C., \& Stathi, S. (2020). \emph{Secure and insecure ingroup
identification as mediators between identity threat and psychological
well-being and intergroup bias} {[}Manuscript in preparation{]}. Faculty
of Arts and Social Sciences, Sabanci University.

\leavevmode\hypertarget{ref-423}{}%
*Bagci, S. C., Stathi, S., \& Piyale, Z. E. (2019). Imagined contact
facilitates acculturation, sometimes: Contradicting evidence from two
sociocultural contexts. \emph{Cultural Diversity and Ethnic Minority
Psychology}, \emph{25}(4), 539--552.
\url{https://doi.org/10.1037/cdp0000256}

\leavevmode\hypertarget{ref-322}{}%
*Bagci, S. C., \& Turnuklu, A. (2019). Intended, unintended, and unknown
consequences of contact: The role of positive-negative contact on
outgroup attitudes, collective action tendencies, and psychological
well-being. \emph{Social Psychology}, \emph{50}(1), 7--23.
\url{https://doi.org/10.1027/1864-9335/a000355}

\leavevmode\hypertarget{ref-414}{}%
*Bagci, S. C., Turnuklu, A., \& Bekmezci, E. (2018a). Cross-group
friendships and psychological well-being: A dual pathway through social
integration and empowerment. \emph{British Journal of Social
Psychology}, \emph{57}(4), 773--792.
\url{https://doi.org/10.1111/bjso.12267}

@@

\leavevmode\hypertarget{ref-2396}{}%
*Bagci, S. C., Verkuyten, M., Koc, Y., Turnuklu, A., Piyale, Z. E., \&
Bekmezci, E. (2020). Being tolerated and being discriminated against:
Links to psychological well-being through threatened social identity
needs. \emph{European Journal of Social Psychology}, \emph{50}(7),
1463--1477. \url{https://doi.org/10.1002/ejsp.2699}

\leavevmode\hypertarget{ref-2075}{}%
*Becker, J. C., Wright, S. C., Lubensky, M. E., \& Zhou, S. (2013).
Friend or ally: Whether cross-group contact undermines collective action
depends on what advantaged group members say (or don't say).
\emph{Personality and Social Psychology Bulletin}, \emph{39}(4),
442--455. \url{https://doi.org/10.1177/0146167213477155}

@@

\leavevmode\hypertarget{ref-2397}{}%
*Bond, J., \& Dixon, J. (2020). \emph{White and Black UK residents'
support for collective action for the removal of public symbols
associated with historical racism in England} {[}Unpublished dataset{]}.
The Open University.

@@

\leavevmode\hypertarget{ref-2398}{}%
*Bracegirdle, C. (2020). \emph{Social network study of two diverse
schools in Northern England} {[}Unpublished dataset{]}. University of
Oxford.

@@

\leavevmode\hypertarget{ref-480}{}%
*Castro, V. S. (2005). Discriminación percibida y autoestima en jóvenes
de grupos minoritarios y mayoritarios en Costa Rica {[}Perceived
discrimination and self-esteem among ethnic majority and minority youths
in Costa Rica{]}. \emph{Revista Interamericana de Psicología},
\emph{39}(1), 93--106.

\leavevmode\hypertarget{ref-1914}{}%
*Castro Solano, A., \& Lupano Perugini, M. L. (2013). Predictores de la
adaptación sociocultural de estudiantes universitarios extranjeros en
Argentina {[}Predictors of socio-cultural adaptation of foreign
university students in Argentina{]}. \emph{Interdisciplinaria},
\emph{30}(2), 265--281.

\leavevmode\hypertarget{ref-3205}{}%
*Cernat, V. (2019). Interminority contact and solidarity under conflict:
Evidence from the Hungarian and Roma minorities in Romania. \emph{Basic
and Applied Social Psychology}, \emph{41}(5), 326--339.
\url{https://doi.org/10.1080/01973533.2019.1652179}

\leavevmode\hypertarget{ref-91}{}%
*Chang, T. K. (2017). \emph{Psychological distress of Asian
international students: Applying intergroup contact theory} {[}Doctoral
dissertation, Indiana University{]}. ProQuest Dissertations and Theses
Global.

\leavevmode\hypertarget{ref-1163}{}%
*Cruse, S. L. (2015). \emph{The relationship of ethnic identity,
perceived discrimination, and outgroup contact to racial attitudes
toward White people in African American college students} {[}Doctoral
dissertation, University of Kansas{]}. ProQuest Dissertations and Theses
Global.

\leavevmode\hypertarget{ref-1248}{}%
*Çakal, H., Hewstone, M., Güler, M., \& Heath, A. (2016). Predicting
support for collective action in the conflict between Turks and Kurds:
Perceived threats as a mediator of intergroup contact and social
identity. \emph{Group Processes \& Intergroup Relations}, \emph{19}(6),
732--752. \url{https://doi.org/10.1177/1368430216641303}

\leavevmode\hypertarget{ref-284}{}%
*Çakal, H., Hewstone, M., Schwär, G., \& Heath, A. (2011). An
investigation of the social identity model of collective action and the
``sedative'' effect of intergroup contact among Black and White students
in South Africa. \emph{British Journal of Social Psychology},
\emph{50}(4), 606--627.
\url{https://doi.org/10.1111/j.2044-8309.2011.02075.x}

\leavevmode\hypertarget{ref-1614}{}%
*Çelebi, E., Verkuyten, M., \& Smyrnioti, N. (2016). Support for Kurdish
language rights in Turkey: The roles of ethnic group, group
identifications, contact, and intergroup perceptions. \emph{Ethnic and
Racial Studies}, \emph{39}(6), 1034--1051.
\url{https://doi.org/10.1080/01419870.2015.1103881}

\leavevmode\hypertarget{ref-783}{}%
*de Vroome, T., Verkuyten, M., \& Martinovic, B. (2014). Host national
identification of immigrants in the Netherlands. \emph{International
Migration Review}, \emph{48}(1), 1--27.
\url{https://doi.org/10.1111/imre.12063}

\leavevmode\hypertarget{ref-3196}{}%
*Di Bernardo, G. A., Vezzali, L., Stathi, S., McKeown Jones, S., Cocco,
V. M., Saguy, T., \& Dixon, J. (2019). Fostering social change among
advantaged and disadvantaged group members: Integrating intergroup
contact and social identity perspectives on collective action.
\emph{Group Processes \& Intergroup Relations}.
\url{https://doi.org/10.1177/1368430219889134}

\leavevmode\hypertarget{ref-1807}{}%
*Di Saint Pierre, F., Martinovic, B., \& de Vroome, T. (2015). Return
wishes of refugees in the Netherlands: The role of integration, host
national identification and perceived discrimination. \emph{Journal of
Ethnic and Migration Studies}, \emph{41}(11), 1836--1857.
\url{https://doi.org/10.1080/1369183X.2015.1023184}

\leavevmode\hypertarget{ref-93}{}%
*Dixon, J., Çakal, H., Khan, W., Osmany, M., Majumdar, S., \& Hassan, M.
(2017). Contact, political solidarity and collective action: An Indian
case study of relations between historically disadvantaged communities.
\emph{Journal of Community \& Applied Social Psychology}, \emph{27}(1),
83--95. \url{https://doi.org/10.1002/casp.2296}

@@

\leavevmode\hypertarget{ref-336}{}%
*Dixon, J., Durrheim, K., Tredoux, C., Tropp, L. R., Clack, B., \&
Eaton, L. (2010). A paradox of integration? Interracial contact,
prejudice reduction, and perceptions of racial discrimination.
\emph{Journal of Social Issues}, \emph{66}(2), 401--416.
\url{https://doi.org/10.1111/j.1540-4560.2010.01652.x}

\leavevmode\hypertarget{ref-dixon_beyond_2012}{}%
Dixon, J., Levine, M., Reicher, S., \& Durrheim, K. (2012). Beyond
prejudice: Are negative evaluations the problem and is getting us to
like one another more the solution? \emph{Behavioral and Brain
Sciences}, \emph{35}(6), 411--425.
\url{https://doi.org/10.1017/S0140525X11002214}

>>>>

\documentclass[twocolumn, 11pt, letterpaper]{article}

% Font
\usepackage[utf8]{inputenc}
\usepackage{MinionPro} 
\input{glyphtounicode}
\pdfgentounicode=1 
\usepackage{microtype}

% Format
\usepackage[letterpaper, margin = 1in]{geometry}
\setlength{\columnsep}{18pt}
\setcounter{secnumdepth}{0}
\usepackage[compact]{titlesec}  
\titleformat*{\section}{\centering\normalfont\Large\bfseries}
\usepackage{authblk}
\renewcommand\Affilfont{\small}
\newcommand{\refsection}{\section{References}}
\providecommand{\tightlist}{\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

% References
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newenvironment{CSLReferences}%
  {}%
  {\par}

% Frontmatter
\title{ Meta-Analysis of the `Ironic' Effects of Intergroup
Contact\thanks{} }
\author[1]{Nils K. Reimer}
\author[2]{Nikhil K. Sengupta}
\affil[1]{University of Southern California, Los Angeles, United States}
\affil[2]{University of Kent, Canterbury, United Kingdom}
\date{June 15, 2021}

@@

\twocolumn[
  \begin{@twocolumnfalse}
    \maketitle
    \begin{abstract}
      \noindent Growing evidence suggests that intergroup contact,
psychology's most-researched paradigm for reducing prejudice, has the
`ironic' effect of reducing support for social change in disadvantaged
groups. We conducted a preregistered meta-analytic test of this effect
across 98 studies with 140 samples of 213,085 disadvantaged-group
members. As predicted, intergroup contact was, on average, associated
with less perceived injustice (\(r = -.07\)), collective action
(\(r = -.06\)), and support for reparative policies (\(r = -.07\)).
However, these associations were small, variable, and consistent with
alternative explanations. Across outcomes, 25--36\% of studies found
\emph{positive} associations with intergroup contact. Moderator analyses
explained about a third of the between-sample variance, showing that, on
average, associations were negative only in studies of adults that
measured intergroup contact directly and were strongest in studies that
examined short-term migration or (post-)colonial intergroup relations.
We also found evidence for an alternative explanation for the apparent
`ironic' effects of intergroup contact as, after controlling for the
positive association of negative contact with support for social change,
positive contact was no longer associated with any of the outcomes. We
close by discussing strengths and limitations of the available evidence
and by highlighting open questions about the relationship between
intergroup contact and support for social change in disadvantaged
groups.
    \end{abstract}
  \end{@twocolumnfalse}
  % \renewcommand{\abstractname}{Statement of Relevance}
  % \begin{@twocolumnfalse}
  %   \maketitle
  %   \begin{abstract}
  %   \noindent 
  %   \end{abstract}
  % \end{@twocolumnfalse}
]

{
  \renewcommand{\thefootnote}%
    {\fnsymbol{footnote}}
  \footnotetext[1]{Manuscript under review.}
}

\noindent It is ironic that psychology's most-researched paradigm for
reducing prejudice---intergroup contact---should hinder, rather than
help, social change. Hundreds of studies have confirmed that intergroup
contact is associated with less negative feelings toward outgroup
members (for a meta-analysis, see Pettigrew \& Tropp, 2006). Most of
these studies have focused on advantaged-group members' contact
experiences with disadvantaged-group members (L. R. Tropp \& Pettigrew,
2005), guided by the idea that reducing advantaged-group members'
prejudice will prevent interpersonal discrimination. Reducing prejudice,
however, is rarely enough to overcome institutional and structural
discrimination. Instead, as the 2020 Black Lives Matter protests have
shown, social change often requires political mobilization by the
disadvantaged. Critics have argued that, by fostering harmony and
reducing conflict, intergroup contact might distract disadvantaged-group
members from the injustice they face and, thereby, undermine their
motivation for social change (Dixon et al., 2005; Reicher, 2007; Wright,
2003). By reducing disadvantaged-group members' opposition to injustice,
intergroup contact might thus have the `ironic' (Saguy et al., 2009),
`sedative' (Çakal et al., 2011), or `paradoxical' (Dixon et al., 2010)
effect of hindering social change.

@@

Collective action is any action directed at improving the conditions of
the disadvantaged ingroup (Wright et al., 1990). Collective action is
crucial to achieving social change because advantaged-group members
rarely give up their advantages without sustained pressure from
disadvantaged-group members (Blumer, 1958; for a review, see Dixon et
al., 2012). Researchers have argued that intergroup contact could
discourage disadvantaged-group members from engaging in collective
action by diminishing their perceived discrimination (Linda R. Tropp et
al., 2012), by blurring boundaries between the disadvantaged ingroup and
the advantaged outgroup (Saguy et al., 2009), by reducing negative
attitudes toward the advantaged outgroup (Wright \& Lubensky, 2009), and
by quelling their anger (Hayward et al., 2018; Tausch et al., 2015). To
the extent that it discourages collective action, intergroup contact
reduces support for social change.

@@

As preregistered, we selected studies in three stages. First, we
screened records based on their title, abstract, and keywords. We
refined our coding strategy over three random samples of 100 records
until we achieved acceptable inter-rater agreement (\(\kappa_1 = .56\),
\(\kappa_2 = .60\), \(\kappa_3 = .79\)). We then divided the remaining
records between the two authors. For each record, one of the authors
coded whether the record met the eligibility criteria (yes, maybe, no),
or whether it was a relevant review article. We kept all records coded
as ``yes'' or ``maybe.'' Second, both authors reviewed each of the
full-text manuscripts from the previous stage and coded whether any
study or sample in the manuscript fulfilled the preregistered
eligibility criteria (\(\kappa = .75\)). Third, we resolved any
disagreements (by consensus) and excluded ineligible records.

@@

Developing methods to detect and correct for publication bias and other
meta-biases is an active area of research, with no single method
outperforming all others (Carter et al., 2019). Following Carter et
al.'s recommendations, we compared results of several methods to adjust
meta-analytic estimates for publication bias: the PET-PEESE estimator
(Stanley \& Doucouliagos, 2014); the three-parameter selection model
{[}3PSM; Vevea \& Hedges (1995){]}; and the \emph{p}-uniform* estimator
(van Aert \& Assen, 2018). In addition, we ran a subgroup analysis that
included only studies that were not published or not conducted with the
intention to examine the effects of intergroup contact on any of the
outcomes---and that, therefore, can be assumed to have not been affected
by publication bias. As these methods do not use Bayesian statistics, we
compared adjusted estimates to unadjusted estimates from a
random-effects meta-analysis estimated with restricted maximum
likelihood estimation using the \emph{metafor} package (Viechtbauer,
2010).

@@

First, the estimated effect sizes for the average associations of
intergroup contact with perceived injustice (\(r = -.07\)), collective
action (\(r = -.06\)), and policy support (\(r = -.07\)) were small.
Across the three outcomes, the 95\% most plausible estimates included
effect sizes between \(r = -.14\) and \(r = .02\). Effect sizes were
thus much smaller than for the association between contact and prejudice
in minority (\(r = -.18\)) and majority (\(r = -.23\)) groups (L. R.
Tropp \& Pettigrew, 2005).\footnote{Comparing standardized effect sizes
  assumes that the two outcomes are of equal importance. That is not
  always the case: for example, a drug that reduces mild symptoms by
  0.20 standard deviations is not better than a drug that reduces deaths
  by 0.10 standard deviations.} Small effects can still be important if
they accumulate over time or across people. For example, even a small
change in policy attitudes could, if it affects enough people, sway a
tight election. Cumulative effects, however, should not be assumed
without an empirical or theoretical rationale (Funder \& Ozer, 2019).

@@

Third, the available evidence is almost entirely from cross-sectional,
observational studies and thus consistent with alternative explanations
for the observed associations. Support for social change could reduce
intergroup contact rather than the other way around. For example,
disadvantaged-group members involved in collective action might avoid
forming friendships with advantaged-group members. Alternatively, the
observed associations could be spurious with both being caused by an
unobserved confounder. For example, a disadvantaged-group member's
socioeconomic status might both expose them to more advantaged-group
members and reduce their perception of injustice. Future research should
prioritize longitudinal studies to confirm the direction of the observed
associations and (field) experiments to rule out confounding and other
alternative explanations.\footnote{Published research includes only
  three longitudinal studies (Koschate et al., 2012; Reimer et al.,
  2017; Linda R. Tropp et al., 2012), all of which conflated
  within-person change and between-person stability (Hamaker et al.,
  2015), two experimental studies with a no-contact control condition
  (Becker et al., 2013; Droogendyk et al., 2016), and two intervention
  studies (Reimer et al., 2021; Shani \& Boehnke, 2017).}

@@

Our research was motivated by the broader question whether intergroup
contact helps or hinders social change. By examining the narrower
question whether intergroup contact diminishes support for social change
in disadvantaged groups, we tested the most prominent hypotheses
explaining how intergroup contact might \emph{hinder} social change. We
did not consider whether intergroup contact might also \emph{help}
social change by reducing prejudice (for critical perspectives, see
Dixon et al., 2012; Jackman, 1994) or by motivating the advantaged to
acknowledge and challenge social inequality (Linda R. Tropp \& Barlow,
2018). Further, we did not consider contact between different
disadvantaged groups, which has been found to increase support for
social change among the disadvantaged (Dixon et al., 2015, 2017).
Therefore, this meta-analysis could not provide a complete answer to the
question whether, on balance, intergroup contact helps or hinders social
change.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-3201}{}%
*Albzour, M., Penic, S., Nasser, R., \& Green, E. G. T. (2019). Support
for {`normalization'} of relations between {Palestinians} and
{Israelis}, and how it relates to contact and resistance in the {West}
{Bank}. \emph{Journal of Social and Political Psychology}, \emph{7}(2),
978--996. \url{https://doi.org/10.5964/jspp.v7i2.877}

\leavevmode\hypertarget{ref-2023}{}%
*Bagci, S. C. (2014). \emph{Cross-ethnic friendships in multiethnic
educational settings: {Consequences} for psychological, academic and
motivational outcomes among young adolescents} {[}Doctoral dissertation,
\{Goldsmiths, University of London\}{]}.
\url{https://doi.org/10.25602/GOLD.00011030}

\leavevmode\hypertarget{ref-2395}{}%
*Bagci, S. C., \& Stathi, S. (2020). \emph{Secure and insecure ingroup
identification as mediators between identity threat and psychological
well-being and intergroup bias} {[}Manuscript in preparation{]}.
{Faculty of Arts and Social Sciences, Sabanci University}.

\leavevmode\hypertarget{ref-423}{}%
*Bagci, S. C., Stathi, S., \& Piyale, Z. E. (2019). Imagined contact
facilitates acculturation, sometimes: {Contradicting} evidence from two
sociocultural contexts. \emph{Cultural Diversity and Ethnic Minority
Psychology}, \emph{25}(4), 539--552.
\url{https://doi.org/10.1037/cdp0000256}

\leavevmode\hypertarget{ref-322}{}%
*Bagci, S. C., \& Turnuklu, A. (2019). Intended, unintended, and unknown
consequences of contact: {The} role of positive-negative contact on
outgroup attitudes, collective action tendencies, and psychological
well-being. \emph{Social Psychology}, \emph{50}(1), 7--23.
\url{https://doi.org/10.1027/1864-9335/a000355}

\leavevmode\hypertarget{ref-414}{}%
*Bagci, S. C., Turnuklu, A., \& Bekmezci, E. (2018a). Cross-group
friendships and psychological well-being: {A} dual pathway through
social integration and empowerment. \emph{British Journal of Social
Psychology}, \emph{57}(4), 773--792.
\url{https://doi.org/10.1111/bjso.12267}

@@

\leavevmode\hypertarget{ref-2396}{}%
*Bagci, S. C., Verkuyten, M., Koc, Y., Turnuklu, A., Piyale, Z. E., \&
Bekmezci, E. (2020). Being tolerated and being discriminated against:
Links to psychological well-being through threatened social identity
needs. \emph{European Journal of Social Psychology}, \emph{50},
1463--1477. \url{https://doi.org/10.1002/ejsp.2699}

\leavevmode\hypertarget{ref-2075}{}%
*Becker, J. C., Wright, S. C., Lubensky, M. E., \& Zhou, S. (2013).
Friend or ally: {Whether} cross-group contact undermines collective
action depends on what advantaged group members say (or don't say).
\emph{Personality and Social Psychology Bulletin}, \emph{39}(4),
442--455. \url{https://doi.org/10.1177/0146167213477155}

@@

\leavevmode\hypertarget{ref-2397}{}%
*Bond, J., \& Dixon, J. (2020). \emph{{White} and {Black} {UK}
residents' support for collective action for the removal of public
symbols associated with historical racism in {England}} {[}Unpublished
dataset{]}. The Open University.

@@

\leavevmode\hypertarget{ref-2398}{}%
*Bracegirdle, C. (2020). \emph{Social network study of two diverse
schools in {Northern England}} {[}Unpublished dataset{]}. University of
Oxford.

@@

\leavevmode\hypertarget{ref-1914}{}%
*Castro Solano, A., \& Lupano Perugini, M. L. (2013). Predictores de la
adaptación sociocultural de estudiantes universitarios extranjeros en
{Argentina} {[}{Predictors} of socio-cultural adaptation of foreign
university students in {Argentina}{]}. \emph{Interdisciplinaria},
\emph{30}(2), 265--281.

\leavevmode\hypertarget{ref-480}{}%
*Castro, V. S. (2005). Discriminación percibida y autoestima en jóvenes
de grupos minoritarios y mayoritarios en {Costa} {Rica} {[}{Perceived}
discrimination and self-esteem among ethnic majority and minority youths
in {Costa} {Rica}{]}. \emph{Revista Interamericana de Psicología},
\emph{39}(1), 93--106.

\leavevmode\hypertarget{ref-3205}{}%
*Cernat, V. (2019). Interminority contact and solidarity under conflict:
{Evidence} from the {Hungarian} and {Roma} minorities in {Romania}.
\emph{Basic and Applied Social Psychology}, \emph{41}(5), 326--339.
\url{https://doi.org/10.1080/01973533.2019.1652179}

\leavevmode\hypertarget{ref-91}{}%
*Chang, T. K. (2017). \emph{Psychological distress of {A}sian
international students: Applying intergroup contact theory} {[}Doctoral
dissertation, \{Indiana University\}{]}. {ProQuest Dissertations and
Theses Global}.

\leavevmode\hypertarget{ref-1163}{}%
*Cruse, S. L. (2015). \emph{The relationship of ethnic identity,
perceived discrimination, and outgroup contact to racial attitudes
toward {White} people in {African} {American} college students}
{[}Doctoral dissertation, \{University of Kansas\}{]}. {ProQuest
Dissertations and Theses Global}.

\leavevmode\hypertarget{ref-1248}{}%
*Çakal, H., Hewstone, M., Güler, M., \& Heath, A. (2016). Predicting
support for collective action in the conflict between {Turks} and
{Kurds}: {Perceived} threats as a mediator of intergroup contact and
social identity. \emph{Group Processes \& Intergroup Relations},
\emph{19}(6), 732--752. \url{https://doi.org/10.1177/1368430216641303}

\leavevmode\hypertarget{ref-284}{}%
*Çakal, H., Hewstone, M., Schwär, G., \& Heath, A. (2011). An
investigation of the social identity model of collective action and the
{`sedative'} effect of intergroup contact among {Black} and {White}
students in {South} {Africa}. \emph{British Journal of Social
Psychology}, \emph{50}(4), 606--627.
\url{https://doi.org/10.1111/j.2044-8309.2011.02075.x}

\leavevmode\hypertarget{ref-1614}{}%
*Çelebi, E., Verkuyten, M., \& Smyrnioti, N. (2016). Support for
{Kurdish} language rights in {Turkey}: The roles of ethnic group, group
identifications, contact, and intergroup perceptions. \emph{Ethnic and
Racial Studies}, \emph{39}(6), 1034--1051.
\url{https://doi.org/10.1080/01419870.2015.1103881}

\leavevmode\hypertarget{ref-783}{}%
*de Vroome, T., Verkuyten, M., \& Martinovic, B. (2014). Host national
identification of immigrants in the {Netherlands}. \emph{International
Migration Review}, \emph{48}(1), 1--27.
\url{https://doi.org/10.1111/imre.12063}

\leavevmode\hypertarget{ref-3196}{}%
*Di Bernardo, G. A., Vezzali, L., Stathi, S., McKeown Jones, S., Cocco,
V. M., Saguy, T., \& Dixon, J. (2019). Fostering social change among
advantaged and disadvantaged group members: {Integrating} intergroup
contact and social identity perspectives on collective action.
\emph{Group Processes \& Intergroup Relations}.
\url{https://doi.org/10.1177/1368430219889134}

\leavevmode\hypertarget{ref-1807}{}%
*Di Saint Pierre, F., Martinovic, B., \& de Vroome, T. (2015). Return
wishes of refugees in the {Netherlands}: {The} role of integration, host
national identification and perceived discrimination. \emph{Journal of
Ethnic and Migration Studies}, \emph{41}(11), 1836--1857.
\url{https://doi.org/10.1080/1369183X.2015.1023184}

\leavevmode\hypertarget{ref-93}{}%
*Dixon, J., Çakal, H., Khan, W., Osmany, M., Majumdar, S., \& Hassan, M.
(2017). Contact, political solidarity and collective action: {An}
{Indian} case study of relations between historically disadvantaged
communities. \emph{Journal of Community \& Applied Social Psychology},
\emph{27}(1), 83--95. \url{https://doi.org/10.1002/casp.2296}

@@

\leavevmode\hypertarget{ref-336}{}%
*Dixon, J., Durrheim, K., Tredoux, C., Tropp, L. R., Clack, B., \&
Eaton, L. (2010). A paradox of integration? {Interracial} contact,
prejudice reduction, and perceptions of racial discrimination.
\emph{Journal of Social Issues}, \emph{66}(2), 401--416.
\url{https://doi.org/10.1111/j.1540-4560.2010.01652.x}

\leavevmode\hypertarget{ref-dixon_beyond_2012}{}%
Dixon, J., Levine, M., Reicher, S., \& Durrheim, K. (2012). Beyond
prejudice: {Are} negative evaluations the problem and is getting us to
like one another more the solution? \emph{Behavioral and Brain
Sciences}, \emph{35}(6), 411--425.
\url{https://doi.org/10.1017/S0140525X11002214}
